Interactive Tools for Science Advice

### 2019-11-04 analysis/presentation/interactive_tools.Rmd
Draft seminar up to NCAM dashboard section

orca(p, file = "analysis/presentation/figures/bg-fig.png", 
     height = 700, width = 1100, scale = 6)

@@

## Background {data-background="figures/bg-fig.png"}

@@

```{r, out.width = "100%"}
knitr::include_graphics("figures/doc_screenshots.PNG")
```

@@

```{r, out.width = "100%"}
# knitr::include_graphics("figures/canada_poll_tracker.PNG")
knitr::include_url("https://newsinteractives.cbc.ca/elections/poll-tracker/canada/", height = "300px")
```

@@

- Using examples we aim to demonstrate that interactive tools can simplify our workflow, from data and model exploration to communication
<br>
- Examples:
    1. **RStrap Explorer:** A tool for examinging fishery-independent survey data
    2. **Tag Explorer:** A mapping tool developed for the exploration of a long-term tagging study
    3. **NCAM Explorer:** A tool for exploring results from the Northern Cod Assessment Model

@@

<br>
<img src="figures/teleost_twitter_crop.jpg" width="100%"/>
<font size="1"> 
<br>
&nbsp; <a href="https://twitter.com/coastguardcan/status/879410397790515201">Canadian Coast Guard | Twitter</a>
</font>

@@

## RStrap Explorer {data-background="figures/RStrap_dashboard_biomass.png" data-background-position="center bottom"}

## RStrap Explorer {data-background="figures/RStrap_dashboard_freq.png" data-background-position="center bottom"}

@@

<img src="figures/cod_tagging.jpg" width="100%"/>
<font size="1"> 
Photo: John Brattey 
</font>

@@

<div style="float: left; width: 25.6%; padding: 1%">
<div class="drop-shadow" style="background-color: rgb(236, 192, 107);">
<img src="figures/taggart_1.png" width="100%"/>
</div>
</div>

@@

<div class="drop-shadow">
<img src="figures/taggart_2.png" width="100%"/>
</div>

@@

<div class="drop-shadow">
<img src="figures/taggart_3.png" width="100%"/>
</div>

@@

- A dashboard was created using the **`shiny`**, **`shinydashboard`**, and **`leaflet`** packages to dynamically explore this large dataset
- Contains
    - One primary page, **Main map**, where release and recovery locations are mapped
    - A series of summary pages where these data are summarised

## Tag Explorer {data-background="figures/tagging_dashboard_map_tab.png" data-background-position="center bottom"}

## Tag Explorer {data-background="figures/tagging_dashboard_summary_tab.png" data-background-position="center bottom"}

>>>>

orca(p, file = "analysis/presentation/graphics/bg-fig.png", 
     height = 700, width = 1100, scale = 6)

@@

## Background {data-background="graphics/bg-fig.png"}

@@

```{r, out.width = "100%"}
knitr::include_graphics("graphics/doc_screenshots.PNG")
```

@@

```{r, out.width = "100%"}
# knitr::include_graphics("graphics/canada_poll_tracker.PNG")
knitr::include_url("https://newsinteractives.cbc.ca/elections/poll-tracker/canada/", height = "300px")
```

@@

- Using examples we aim to demonstrate that interactive tools can simplify our workflow, from data and model exploration to communication
<br>
- Examples:
    1. **RStrap Explorer:** A tool for examining fishery-independent survey data
    2. **Tag Explorer:** A mapping tool developed for the exploration of a long-term tagging study
    3. **NCAM Explorer:** A tool for exploring results from the Northern Cod Assessment Model

@@

<br>
<img src="graphics/teleost_twitter_crop.jpg" width="100%"/>
<font size="1"> 
<br>
&nbsp; <a href="https://twitter.com/coastguardcan/status/879410397790515201">Canadian Coast Guard | Twitter</a>
</font>

@@

## RStrap Explorer {data-background="graphics/RStrap_dashboard_biomass.png" data-background-position="center bottom"}

## RStrap Explorer {data-background="graphics/RStrap_dashboard_freq.png" data-background-position="center bottom"}

@@

<img src="graphics/cod_tagging.jpg" width="100%"/>
<font size="1"> 
Photo: John Brattey 
</font>

@@

<div style="float: left; width: 25.6%; padding: 1%">
<div class="drop-shadow" style="background-color: rgb(236, 192, 107);">
<img src="graphics/taggart_1.png" width="100%"/>
</div>
</div>

@@

<div class="drop-shadow">
<img src="graphics/taggart_2.png" width="100%"/>
</div>

@@

<div class="drop-shadow">
<img src="graphics/taggart_3.png" width="100%"/>
</div>

@@

- A dashboard was created using the **`shiny`**, **`shinydashboard`**, and **`leaflet`** packages to dynamically explore this large data-set
- Contains
    - One primary page, **Main map**, where release and recovery locations are mapped
    - A series of summary pages where these data are summarized

## Tag Explorer {data-background="graphics/tagging_dashboard_map_tab.png" data-background-position="center bottom"}

## Tag Explorer {data-background="graphics/tagging_dashboard_summary_tab.png" data-background-position="center bottom"}

## NCAM Explorer

<div style="float: left; width: 50%; padding-right: 5%;">

- Integrated stock assessment models are becoming more common
- Advances in computational power and methods $\rightarrow$ possibilities
- Requires a solid understanding of the inputs, methods and outputs
    - Understanding is a prerequisite for communication

<div class="drop-shadow" style="float: right; width: 40%; padding: 2.5%; font-size: 8px;">

**Process equations**  
<br>
Stochastic cohort model with a plus group to model the unobserved states:
$$\log(N_{a,y}) = \left\{\begin{matrix}
 \log(N_{a-1,y-1}) - Z_{a-1,y-1} + \delta_{a,y}, & a < A \\
 \log\{N_{a-1,y-1}\exp(-Z_{a-1,y-1}) + N_{a,y-1}\exp(-Z_{a,y-1})\} + \delta_{a,y}, & a = A.
\end{matrix}\right. $$

The ages are 1-10+ and years are 1975-2015. $Z_{a,y} = F_{a,y} + M_{a,y}$, where $M_{a,y} = 0.2$ is the base case assumption.

Recruitments, $N_{1,1}, ... N_{1,Y}$, are treated as uncorrelated lognormal random variables
$$\log(N_{1,y}) \overset{iid}\sim N(r, \sigma_{r}^2).$$

Catches are modeled using the Baranov catch equation,
$$C_{a,y} = N_{a,y}\{1 - \exp(-Z_{a,y})\}F_{a,y}/Z_{a,y}.$$

Fishing moralities are modeled as a stochastic process, with
$$Cov\{\log(F_{a,y}),\log(F_{a-j,y-k})\} = \frac{\sigma_{F}^2 \varphi_{F,a}^j \varphi_{F,y}^k}{(1-\varphi_{F,a}^2)(1-\varphi_{F,y}^2)}.$$

**Observation equations**  
<br>
The model predicted catch for survey $s$ is
$$\log(I_{s,a,y}) = \log(q_{s,a}) + \log(N_{a,y}) - t_{s,y}Z_{a,y} + \varepsilon_{s,a,y}, ~~ \varepsilon_{s,a,y} \overset{iid}\sim N(0, \sigma_{s,G(a)}^2).$$

Survey variance was split out and self-weighted by age groups 1-3, 4-7, and 8-10+

Total catch and age compositions were treated separately. Total catch was modeled as lognormal, 
$$\log(C_{obs,y}) = \log(C_{y}) + \varepsilon_{C,y}, ~~ \varepsilon_{C,y} \overset{iid}\sim N(0, \sigma_{C}^2)$$
Age compositions were modeled as multiplicative logistic normal with a censored component for zero's

## NCAM Explorer <img src="graphics/cod.png" height="100px" style="margin-bottom: -35px; margin-left: 10px;">

- **Challenge:** the data-rich case of Northern cod
    - Assessed using an age-structured, state-space assessment model, called NCAM, which integrates:
        - Research vessel autumn trawl surveys (1983-present)
        - Sentinel fishery surveys (1995-present)
        - Inshore acoustic surveys (1995-2009)
        - Fishery catch-at-age compositions and partial fishery landings (1983-present)
        - Tagging data (1983-present)

## NCAM Explorer

> - **Communication:** standard approach involves compiling and presenting static documentation

<div style="float: left; width: 20%; text-align: center;">
Code  
<font size="20"><i class="fa fa-file-code"></i></font>
</div>

<div style="float: left; width: 20%; text-align: center;">
Plots & Tables  
<font size="20"><i class="fa fa-file-image"></i></font>
<font size="20"><i class="fa fa-file-excel"></i></font>
</div>

<div style="float: left; width: 20%; text-align: center;">
Documentation  
<font size="20"><i class="fa fa-file-word"></i></font>
<font size="20"><i class="fa fa-file-powerpoint"></i></font>
</div>

- Involves a lot of copy-pasting
- Important details may not be visible
- Linear format may be the best way to explore ideas

## NCAM Explorer

> - A dashboard was created using the **`flexdashboard`**, **`plotly`**, and **`crosstalk`** packages to explore and communicate results from NCAM

<div style="float: left; width: 33.3333%; text-align: center;">
Code  
<font size="20"><i class="fa fa-file-code"></i></font>
</div>

<div style="float: left; width: 33.3333%; text-align: center;">
<br>
$\rightarrow$
</div>

<div style="float: left; width: 33.3333%; text-align: center;">
HTML  
<font size="20"><i class="far fa-file-code"></i></font>
</div>

- Contains a series of pages with plots and tables typically presented at assessment meetings

## NCAM Explorer

<iframe src="graphics/NCAM_dashboard.html" style="height: 700px; width: 1100px; margin-top: -150px; margin-left: -60px; margin-right: -60px; position: absolute; z-index: 1;"></iframe>
